---
title: "Data 607 Assignment Week 5"
author: "Ajay Arora"
date: "September 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#1. Import libraries
```{r}
library(stringr)
library(tidyr)
library(dplyr)
library('scales')
library(ggplot2)
library(kableExtra)
```

#2. Get .csv file

```{r}
csvfile <- read.csv("https://raw.githubusercontent.com/AjayArora35/Data-607-Assignment-Week-5/master/AssignmentWeek5.csv", header=FALSE, sep=",", stringsAsFactors=FALSE)
csvfile <- data.frame(csvfile)
csvfile
```

#3. Assign headers to 'V1' and 'V2' then copy the first row as headers, etc.

```{r}
#Assign missing headers
csvfile$V1[1] <- "Airline"
csvfile$V2[1] <- "Status"
csvfile
#Copy the first row as headers
names(csvfile) <- csvfile[1,]
csvfile
#Remove the first row
csvfile <- csvfile[-c(1), ]
csvfile
#Remove the NA row as well
csvfile <- csvfile[c(1,2,4,5), ]
csvfile

#Fill in the Airline missing column data 
for (i in 1:dim(csvfile)[1]){
  if (i %% 2 == 0){
      csvfile$Airline[i] <-  csvfile$Airline[i-1]
  }
}


csvfile
```
#4. Some Analysis

#Now, take the wide table and make it long: tidyr (Pivot), and additional cleanup

```{r}
flights <- gather(csvfile, key = 'City', value = 'Flights', 3:7, convert=TRUE)
flights

#Remove new line character from city names and remove commas from number of flights
flights$City <- str_replace_all(flights$City, "[\r\n]" , "")
flights

flights$Flights <- str_replace_all(flights$Flights, "[,]" , "")
flights

#Now, separate out the status column into On-time and Delayed 
flights2 <-  flights %>% spread(Status, `Flights`, convert=TRUE)
flights2

#Now, add some percentages for On-Time and Delayed
flights2 <- flights2 %>% 
  mutate(total = delayed + `on time`,
         OnTime = percent(`on time`/total*100, accuracy = .01, scale=1),
         Delayed = percent(delayed/total*100, accuracy = .01, scale=1))
flights2 %>% 
  kable() %>%
  kable_styling()

#Reference: http://www.sthda.com/english/wiki/ggplot2-barplots-quick-start-guide-r-software-and-data-visualization
#Now, graph total flights for each airline
ggplot(flights2,
       aes(x=flights2$Airline,
           y=flights2$total
           ))+
    geom_bar(stat="identity")

#Graph Total flights for each city
ggplot(flights2,
       aes(x=flights2$City,
           y=flights2$total, fill = flights2$Airline
           ))+
    geom_bar(stat="identity", position="dodge")

#Graph percentage delays for all cities per airline
ggplot(flights2,
       aes(x=flights2$City,
           y=flights2$Delayed, fill = flights2$Airline
           ))+
    geom_bar(stat="identity", position="dodge")

```
#5. Final Analysis

```{r}
#Best OnTime Percentage Rate

flights3 <- flights2 %>% filter(flights2$OnTime == max(flights2$OnTime))
flights3%>% 
  kable() %>%
  kable_styling()

#Airline with the best OnTime arrival rate

flights4 <- flights2 %>% 
  group_by(Airline) %>% 
  summarise(Delayed_Flights = sum(delayed), Total_Flights = sum(total)) %>% 
  mutate(Delayed = (Delayed_Flights / Total_Flights)*100)

flights4%>% 
  kable() %>%
  kable_styling()

#Airline delay as percentage
ggplot(flights4,
       aes(x=flights4$Airline,
           y=flights4$Delayed, fill = flights4$Airline
           ))+
    geom_bar(stat="identity", position="dodge")

#Airline Total Flights
ggplot(flights4,
       aes(x=flights4$Airline,
           y=flights4$Total_Flights, fill = flights4$Airline
           ))+
    geom_bar(stat="identity", position="dodge")

#Airline Total Delayed Flights
ggplot(flights4,
       aes(x=flights4$Airline,
           y=flights4$Delayed_Flights, fill = flights4$Airline
           ))+
    geom_bar(stat="identity", position="dodge")
```

#6. Conclusions
As indicated by the above graphs, AM West had many more flights than Alaska.  The best arrival time was by Alaska to Phoneix where their on-time is a 94.85%.  Alaska's overall delay is at 13.27% compared with 10.89% with AM West.  Lastly, for each city, AM West has a larger percentage delay than Alaska.  


